---
title: 'Homework #4'
author: "Katherine Evers"
date: "4/13/2020"
output: html_document
---
Overview
In this homework assignment, you will explore, analyze and model a data set containing approximately 8000
records representing a customer at an auto insurance company. Each record has two response variables. The
first response variable, TARGET_FLAG, is a 1 or a 0. A “1” means that the person was in a car crash. A zero
means that the person was not in a car crash. The second response variable is TARGET_AMT. This value is zero
if the person did not `crash` their car. But if they did crash their car, this number will be a value greater than zero.
Your objective is to build multiple linear regression and binary logistic regression models on the training data
to predict the probability that a person will crash their car and also the amount of money it will cost if the person
does crash their car. You can only use the variables given to you (or variables that you derive from the variables
provided). Below is a short description of the variables of interest in the data set:

```{r packages, echo = FALSE, warning = FALSE, message = FALSE}
library(ggplot2)
library(dplyr)
library(stringr)
library(corrplot)
library(reshape2)
library(tidyr)
library(caret)
library(pROC)
library(gridExtra)
library(pander)
library(mice)
library(car)
```

1. DATA EXPLORATION (25 Points)

Describe the size and the variables in the insurance training data set. Consider that too much detail will cause a
manager to lose interest while too little detail will make the manager consider that you aren’t doing your job. Some
suggestions are given below. Please do NOT treat this as a check list of things to do to complete the assignment.
You should have your own thoughts on what to tell the boss. These are just ideas.
a. Mean / Standard Deviation / Median
b. Bar Chart or Box Plot of the data
c. Is the data correlated to the target variable (or to other variables?)
d. Are any of the variables missing and need to be imputed “fixed”?


```{r, warning = FALSE, message = FALSE}
raw_data <- read.csv("insurance_training_data.csv", header = TRUE)

dim(raw_data)

head(raw_data, 10)

#clean data
training_data <- training_data %>% mutate(
    TARGET_FLAG = as.factor(TARGET_FLAG),
    REVOKED = as.factor(REVOKED),
    INCOME = as.numeric(INCOME),
    HOME_VAL = as.numeric(HOME_VAL),
    BLUEBOOK = as.numeric(BLUEBOOK),
    OLDCLAIM = as.numeric(OLDCLAIM),
    MSTATUS = as.factor(str_remove(MSTATUS, "^z_")),
    SEX = as.factor(str_remove(SEX, "^z_")),
    EDUCATION = as.factor(str_remove(EDUCATION, "^z_")),
    JOB = as.factor(str_remove(JOB, "^z_")),
    CAR_TYPE = as.factor(str_remove(CAR_TYPE, "^z_")),
    URBANICITY = as.factor(str_remove(URBANICITY, "^z_")))

#summary stats
sum_data <- summary(training_data)
pander(sum_data, split.table = 100, style = 'rmarkdown')

#missing values
#training_data %>%
  summarise_all(list(~sum(is.na(.))))

missing.values <- training_data %>%
    gather(key = "key", value = "val") %>%
    mutate(is.missing = is.na(val)) %>%
    group_by(key, is.missing) %>%
    summarise(num.missing = n()) %>%
    filter(is.missing==T) %>%
    select(-is.missing) %>%
    arrange(desc(num.missing)) 

510/8161 #0.06249234
454/8161 #0.05563044
6/8161 #0.000735204

missing.values <- training_data %>%
  gather(key = "key", value = "val") %>%
  mutate(isna = is.na(val)) %>%
  group_by(key) %>%
  mutate(total = n()) %>%
  group_by(key, total, isna) %>%
  summarise(num.isna = n()) %>%
  mutate(pct = num.isna / total * 100)


levels <-
    (missing.values  %>% filter(isna == T) %>% arrange(desc(pct)))$key

percentage.plot <- missing.values %>%
      ggplot() +
        geom_bar(aes(x = reorder(key, desc(pct)), 
                     y = pct, fill=isna), 
                 stat = 'identity', alpha=0.8) +
      scale_x_discrete(limits = levels) +
      scale_fill_manual(name = "", 
                        values = c('steelblue', 'tomato3'), labels = c("Present", "Missing")) +
      coord_flip() +
      labs(title = "Percentage of Missing Values", x =
             'Variable', y = "% of missing values")

percentage.plot

#correlation matrix
cor_data <- data.frame(lapply(training_data, function(x) as.numeric(as.factor(x))))
corrplot(cor(dplyr::select(drop_na(cor_data), everything())), type = "lower")


#histograms
histogram<-training_data[,c(2:26)]
histogram_data <- melt(histogram)
ggplot(histogram_data, aes(value)) + 
  geom_bar(fill = "blue") + 
  facet_wrap(~variable, scales = "free") + 
  geom_histogram(bins=50)


#box plots
plot1 <- ggplot(training_data, aes(x = TARGET_FLAG, y = TARGET_AMT, group=TARGET_FLAG)) + geom_boxplot(aes(fill = TARGET_FLAG)) + coord_flip()
plot2 <- ggplot(training_data, aes(x = TARGET_FLAG, y = AGE, group=TARGET_FLAG)) + geom_boxplot(aes(fill = TARGET_FLAG)) + coord_flip()
plot3 <- ggplot(training_data, aes(x = TARGET_FLAG, y = HOMEKIDS, group=TARGET_FLAG)) + geom_boxplot(aes(fill = TARGET_FLAG)) + coord_flip()
plot4 <- ggplot(training_data, aes(x = TARGET_FLAG, y = YOJ, group=TARGET_FLAG)) + geom_boxplot(aes(fill = TARGET_FLAG)) + coord_flip()
plot5 <- ggplot(training_data, aes(x = TARGET_FLAG, y = INCOME, group=TARGET_FLAG)) + geom_boxplot(aes(fill = TARGET_FLAG)) + coord_flip()
plot6 <- ggplot(training_data, aes(x = TARGET_FLAG, y = PARENT1, group=TARGET_FLAG)) + geom_boxplot(aes(fill = TARGET_FLAG)) + coord_flip()
plot7 <- ggplot(training_data, aes(x = TARGET_FLAG, y = HOME_VAL, group=TARGET_FLAG)) + geom_boxplot(aes(fill = TARGET_FLAG)) + coord_flip()
plot8 <- ggplot(training_data, aes(x = TARGET_FLAG, y = TRAVTIME, group=TARGET_FLAG)) + geom_boxplot(aes(fill = TARGET_FLAG)) + coord_flip()
plot9 <- ggplot(training_data, aes(x = TARGET_FLAG, y = BLUEBOOK, group=TARGET_FLAG)) + geom_boxplot(aes(fill = TARGET_FLAG)) + coord_flip()
plot10 <- ggplot(training_data, aes(x = TARGET_FLAG, y = TIF, group=TARGET_FLAG)) + geom_boxplot(aes(fill = TARGET_FLAG)) + coord_flip()
plot11 <- ggplot(training_data, aes(x = TARGET_FLAG, y = OLDCLAIM, group=TARGET_FLAG)) + geom_boxplot(aes(fill = TARGET_FLAG)) + coord_flip()
plot12 <- ggplot(training_data, aes(x = TARGET_FLAG, y = CLM_FREQ, group=TARGET_FLAG)) + geom_boxplot(aes(fill = TARGET_FLAG)) + coord_flip()
plot13 <- ggplot(training_data, aes(x = TARGET_FLAG, y = MVR_PTS, group=TARGET_FLAG)) + geom_boxplot(aes(fill = TARGET_FLAG)) + coord_flip()
plot14 <- ggplot(training_data, aes(x = TARGET_FLAG, y = CAR_AGE, group=TARGET_FLAG)) + geom_boxplot(aes(fill = TARGET_FLAG)) + coord_flip()

#grid.arrange(plot1,plot2,plot3,plot4,plot5,plot6,plot7,plot8,plot9,plot10,plot11,plot12, plot13, plot14, ncol=3)

#box plots
distribution <- training_data %>% 
  dplyr::select(c("TARGET_FLAG", "AGE", "HOMEKIDS","YOJ", "INCOME", "HOME_VAL", "TRAVTIME", "BLUEBOOK", "TIF", "OLDCLAIM", "CLM_FREQ", "MVR_PTS", "CAR_AGE")) %>% 
  gather(key, value, -TARGET_FLAG) %>% 
  mutate(value = as.integer(value),
         key = as.factor(key),
         TARGET_FLAG = as.factor(TARGET_FLAG))

distribution  %>% 
  ggplot(aes(x = key, y = value)) +
  geom_boxplot(aes(fill = TARGET_FLAG)) +
  facet_wrap(~ key, scales = 'free', ncol = 3) +
  theme_minimal() +
  coord_flip()


```

```{r}
mlr <- subset(training_data, TARGET_FLAG=='1')
summary(mlr)

#histograms
histogram<-mlr[,c(2:26)]
histogram_data <- melt(histogram)
ggplot(histogram_data, aes(value)) + 
  geom_bar(fill = "blue") + 
  facet_wrap(~variable, scales = "free") + 
  geom_histogram(bins=50)
```

2. DATA PREPARATION (25 Points)

Describe how you have transformed the data by changing the original variables or creating new variables. If you
did transform the data or create new variables, discuss why you did this. Here are some possible transformations.
a. Fix missing values (maybe with a Mean or Median value)
b. Create flags to suggest if a variable was missing
c. Transform data by putting it into buckets
d. Mathematical transforms such as log or square root (or use Box-Cox)
e. Combine variables (such as ratios or adding or multiplying) to create new variables

```{r}
init = mice(training_data) 
meth = init$method
predM = init$predictorMatrix

#predM[c("ID", "TARGET_FLAG")]=0

set.seed(100)
imputed = mice(training_data, method=meth, predictorMatrix=predM, m=5)
imputed <- complete(imputed)
imputed_data <- imputed[,c(2:26)]

missing <- imputed_data[,c(4,6,24)]
summary(missing)

#impute2 <- mice(training_data, method = 'rf', predictorMatrix=predM)
#imputed2 <- complete(impute2)

#log transform target amt
```

```{r}
transformation <- imputed %>% mutate_at(c("TARGET_AMT", "KIDSDRIV", "HOMEKIDS", "TIF", "CLM_FREQ", "MVR_PTS"), log)

#histograms
histogram<-transformation[,c(2:26)]
histogram_data <- melt(histogram)
ggplot(histogram_data, aes(value)) + 
  geom_bar(fill = "blue") + 
  facet_wrap(~variable, scales = "free") + 
  geom_histogram(bins=50)
```

3. BUILD MODELS (25 Points)

Using the training data set, build at least two different multiple linear regression models and three different binary
logistic regression models, using different variables (or the same variables with different transformations). You
may select the variables manually, use an approach such as Forward or Stepwise, use a different approach such
as trees, or use a combination of techniques. Describe the techniques you used. If you manually selected a
variable for inclusion into the model or exclusion into the model, indicate why this was done.
Discuss the coefficients in the models, do they make sense? For example, if a person has a lot of traffic tickets,
you would reasonably expect that person to have more car crashes. If the coefficient is negative (suggesting that
the person is a safer driver), then that needs to be discussed. Are you keeping the model even though it is counter
intuitive? Why? The boss needs to know.

```{r}
#logistic
#logit_data <- data.frame(lapply(imputed_data, function(x) as.numeric(as.factor(x)))) %>% 
  #mutate(TARGET_FLAG = as.factor(TARGET_FLAG))

logit_data <- imputed_data[,c(1, 3:25)]

model <- glm(TARGET_FLAG ~ KIDSDRIV + MSTATUS + SEX + EDUCATION + JOB + TRAVTIME + CAR_USE + CAR_TYPE + OLDCLAIM + CLM_FREQ + CAR_AGE + URBANICITY, family = "binomial", logit_data)
summary(model)

#test
confusionMatrix(as.factor(as.integer(fitted(model) > .5)), as.factor(model$y), positive = "1")

roc(logit_data$TARGET_FLAG,  predict(model, logit_data, interval = "prediction"))

#metrics1 <- c(c1$overall[1], "Class. Error Rate" = 1 - as.numeric(c1$overall[1]), c1$byClass[c(1, 2, 5, 7)], AUC = roc1$auc)

plot(roc(logit_data$TARGET_FLAG,  predict(model, logit_data, interval = "prediction")), print.auc = TRUE, col="red", main = "ROC Curve")


###


target <- logit_data %>% select(-TARGET_FLAG)

test_results <- predict(model, newdata = target, type = "response")

results <- bind_cols(logit_data , data.frame(scored_target_flag = test_results)) %>% 
    mutate(scored_target_flag = if_else(scored_target_flag > 0.5, 1, 0)) #%>% print

cm <- confusionMatrix(as.factor(results$scored_target_flag), as.factor(results$target_flag), positive = "1", mode = "everything") %>% print 


curveRoc <- roc(results$target, results$scored_target)
curveRoc$auc



plot(curveRoc[[1]], 
     main = "ROC Curve",
     xlab = "False Positive Rate",
     ylab = "True Positive Rate")
```

```{r}
#mlr models
#remove target_flag column
mlr_data <- imputed_data[,c(2:25)]
mlr_data2 <- transformation[,c(2:25)]

#model 1
model1 <- lm(TARGET_AMT ~ ., mlr_data)
summary(model1)

vif(model1)

par(mfrow=c(2,2))
plot(model1)


model1_2 <- lm(TARGET_AMT ~ ., mlr_data2)
summary(model1)

vif(model1_2)

par(mfrow=c(2,2))
plot(model1_2)

#model 2
model2 <- lm(TARGET_AMT ~ KIDSDRIV + PARENT1 + HOME_VAL + MSTATUS  + TRAVTIME + CAR_USE + TIF + CAR_TYPE + CLM_FREQ + REVOKED + MVR_PTS  + CAR_AGE + URBANICITY, mlr_data)
summary(model2)

vif(model2)

par(mfrow=c(2,2))
plot(model2)


model2_1 <- lm(TARGET_AMT ~ KIDSDRIV + PARENT1 + HOME_VAL + MSTATUS  + TRAVTIME + CAR_USE + TIF + CAR_TYPE + CLM_FREQ + REVOKED + MVR_PTS  + CAR_AGE + URBANICITY, mlr_data2)
summary(model2_1)

vif(model2)

par(mfrow=c(2,2))
plot(model2)

model3 <- lm(TARGET_AMT ~ KIDSDRIV + HOMEKIDS + INCOME + PARENT1 + HOME_VAL + MSTATUS + EDUCATION + TRAVTIME + CAR_USE + BLUEBOOK + TIF + CAR_TYPE + CLM_FREQ + REVOKED + MVR_PTS + CAR_AGE + URBANICITY, mlr_data)
summary(model3)

null.model <- lm(TARGET_AMT ~ 1 , data= mlr_data)  # base intercept only model
full.model <- lm(TARGET_AMT ~ . , data= mlr_data)  # full model with all predictors

# perform step-wise selection
model <- step(null.model, scope = list(lower = null.model, upper = full.model), direction = "both", trace = 0, steps = 1000) 

summary(model)

vif(model)

par(mfrow=c(2,2))
plot(model)
```
model_1 <- glm(target ~ indus + nox + age + dis + rad + tax, family = binomial(link = "logit"), data = training_data)


4. SELECT MODELS (25 Points)

Decide on the criteria for selecting the best multiple linear regression model and the best binary logistic regression model. Will you select models with slightly worse performance if it makes more sense or is more parsimonious? Discuss why you selected your models. For the multiple linear regression model, will you use a metric such as Adjusted R2, RMSE, etc.? Be sure to explain how you can make inferences from the model, discuss multi-collinearity issues (if any), and discuss other relevant model output. Using the training data set, evaluate the multiple linear regression model based on (a) mean squared error, (b) R2, (c) F-statistic, and (d) residual plots. For the binary logistic regression model, will you use a metric such as log likelihood, AIC, ROC curve, etc.? Using the training data set, evaluate the binary logistic regression model based on (a) accuracy, (b) classification error rate, (c) precision, (d) sensitivity, (e) specificity, (f) F1 score, (g) AUC, and (h) confusion matrix. Make predictions using the evaluation data set